\section{Preliminaries: Halal Trust Context and Ethical Design Requirements}
\label{sec:prelim}

This section provides contextual background on halal certification trust gaps and derives \emph{engineering design requirements} from Islamic ethical principles.
It does not constitute a religious ruling (fatwa); rather, it informs system design analogously to value-sensitive design, where moral considerations shape technical choices without replacing domain expertise~\cite{friedman2017value}.

\subsection{Halal Certification Context in Indonesia}
Indonesia's halal certification ecosystem involves BPJPH registration, product classification, and MUI auditing for many categories~\cite{sidarto2021halal,bpjph2022regulation}.
Producers submit ingredient declarations, process descriptions, and supporting laboratory or supplier attestations.
Upon approval, halal labels may be applied to specific products or production lines; export markets increasingly request traceability evidence beyond the label itself~\cite{marianingsih2024halal}.

Reported issues include certificate forgery, relabeling of non-certified batches, and inability of consumers to independently verify batch-level claims at point of purchase~\cite{rashid2018halal,kshetri2018blockchain}.
UMKM producers---small food manufacturers, home industries scaling up, and regional cooperatives---often lack IT staff to operate enterprise traceability suites.
HalalChain therefore targets \emph{batch-level} attestation with minimal on-chain fields and a mobile-friendly consumer verify page.

\subsection{Actors and Trust Boundaries}
Four actor classes interact with the protocol:
\begin{itemize}
  \item \textbf{Producer (UMKM):} Registers batches and uploads evidence to IPFS; cannot self-verify halal status.
  \item \textbf{Auditor:} Reviews evidence and transitions batch status to Verified, Rejected, or Revoked; actions are signed on-chain.
  \item \textbf{Consumer:} Reads public batch status via QR-linked URL; does not hold keys or tokens.
  \item \textbf{Administrator:} Grants \texttt{PRODUCER\_ROLE} and \texttt{AUDITOR\_ROLE}; a trust bottleneck discussed in Section~\ref{sec:discussion}.
\end{itemize}
The protocol does \emph{not} automatically validate ingredient chemistry or slaughter method; it records \emph{which auditor} approved \emph{which documents} at \emph{which time}, shifting trust from opaque databases to inspectable records.

\subsection{From Principles to Design Requirements}
Islamic ethical principles inform \emph{what} the system should optimize for, expressed as testable properties.
We select four principles frequently cited in halal supply chain literature and map them to HalalChain mechanisms in Table~\ref{tab:ethics}.

\begin{table}[t]
  \caption{Mapping of ethical design requirements to HalalChain implementation.}
  \label{tab:ethics}
  \centering
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \begin{tabularx}{\columnwidth}{@{}>{\raggedright\arraybackslash}p{0.24\columnwidth}>{\raggedright\arraybackslash}p{0.26\columnwidth}X@{}}
    \toprule
    \textbf{Principle} & \textbf{Requirement} & \textbf{Implementation} \\
    \midrule
    Amanah (trustworthiness) & Immutable audit record & Append-only status transitions; rejected batches cannot be re-verified \\
    Thoyyib (wholesomeness) & Verified-only halal display & Consumer UI shows certification only if \texttt{status = Verified} \\
    Adl (justice) & Distributed data access & Public chain reads + IPFS gateways; no single vendor gate for status \\
    Hisba (accountability) & Auditor attribution & \texttt{auditor} address, \texttt{auditedAt}, optional \texttt{auditIpfsCid} \\
    \bottomrule
  \end{tabularx}
\end{table}

\emph{Amanah} is operationalized by smart contract guards: once \texttt{rejectBatch} executes, \texttt{verifyBatch} on the same \texttt{batchId} reverts with \texttt{InvalidStatus}, preventing retroactive ``approval'' of a failed batch.
\emph{Thoyyib} appears in the consumer dashboard, which renders a positive halal indicator only for Verified batches---Pending and Rejected states show explicit non-certified messaging in English and Indonesian.
\emph{Adl} is partially satisfied by public read access but constrained because role assignment remains centralized under \texttt{DEFAULT\_ADMIN\_ROLE}.
\emph{Hisba} binds each decision to an Ethereum address and block timestamp, enabling post-hoc accountability if an auditor key is known.

The current prototype grants \texttt{AUDITOR\_ROLE} via an administrator (OpenZeppelin \texttt{DEFAULT\_ADMIN\_ROLE}), which partially limits \emph{Adl}; Section~\ref{sec:discussion} discusses multi-signature and DAO-based mitigations~\cite{gnosis2024safe}.

\FloatBarrier
