\section{Implementation}
\label{sec:implementation}

Table~\ref{tab:stack} summarizes the technology stack.
The open-source repository is available at \url{https://github.com/Fatihmaull/halal-chain}.

\begin{table}[t]
  \caption{HalalChain implementation stack.}
  \label{tab:stack}
  \centering
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \begin{tabularx}{\columnwidth}{@{}>{\raggedright\arraybackslash}p{0.22\columnwidth}X@{}}
    \toprule
    \textbf{Layer} & \textbf{Technology} \\
    \midrule
    Blockchain & Solidity 0.8.20, OpenZeppelin 5.x, Base Sepolia (84532) \\
    Development & Hardhat 3.7, Mocha/Chai (7 tests), TypeScript scripts \\
    Frontend & Next.js 16, Wagmi v3, Viem 2.x, Tailwind CSS \\
    Storage & IPFS via Pinata (\texttt{POST /api/ipfs/upload}) \\
    Deployment & \makecell[tl]{Hardhat (\texttt{deploy:local},\\\texttt{deploy:base-sepolia}), Vercel} \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Smart Contracts}
The \texttt{HalalChain} contract extends OpenZeppelin \texttt{AccessControl}~\cite{openzeppelin2024access}, replacing an earlier single-owner auditor mapping with explicit roles grantable by \texttt{DEFAULT\_ADMIN\_ROLE}.
The constructor assigns admin to the deployer; deployment scripts optionally grant producer and auditor roles to configured addresses for testnet demos.

Unit tests (\texttt{contracts/test/HalalChain.test.ts}) cover:
\begin{itemize}
  \item batch registration and monotonic \texttt{nextBatchId} assignment;
  \item verification and rejection with role enforcement;
  \item revision after rejection with \texttt{NotOriginalProducer} and \texttt{ParentNotRejected} guards;
  \item revocation from Verified state;
  \item unauthorized calls from addresses lacking \texttt{PRODUCER\_ROLE} or \texttt{AUDITOR\_ROLE};
  \item invalid \texttt{batchId} lookups reverting with \texttt{InvalidBatch}.
\end{itemize}

Hardhat scripts include \texttt{deploy.ts} (local or Base Sepolia), \texttt{export-abi.ts} (ABI copy to \texttt{web/src/contracts/}), and \texttt{evaluate-gas.ts} (receipt gas aggregation for evaluation scenarios).

\subsection{Web Application}
Three role-specific dashboards share a common layout and bilingual (English/Indonesian) strings:

\textbf{Producer dashboard} guides users through product naming, multi-file upload, IPFS pinning, and \texttt{registerBatch} submission.
After confirmation, the UI displays the assigned \texttt{batchId}, transaction hash, and a QR code pointing to the consumer verify URL.
Batch history lists prior registrations with status badges and links to parent batches when applicable.

\textbf{Auditor dashboard} presents a pending queue fetched via \texttt{getBatch} iteration or event logs.
Auditors open IPFS document links in new tabs, then submit \texttt{verifyBatch}, \texttt{rejectBatch} (with reason text), or \texttt{revokeBatch} through wallet-connected writes.
Confirmation modals reduce accidental state changes.

\textbf{Consumer verify page} (\texttt{/verify/[batchId]}) performs read-only RPC calls.
It renders green ``Halal Certified'' / ``Tersertifikasi Halal'' for Verified batches, red rejection messaging with \texttt{rejectReason} for Rejected batches, and neutral pending/revoked states otherwise.
Parent batch links enable traceability across revision chains.

\subsection{IPFS Upload Pipeline}
A Next.js API route (\texttt{POST /api/ipfs/upload}) accepts \texttt{multipart/form-data} file uploads, pins via Pinata JWT server-side, and returns the CID JSON payload to the client.
Environment variable \texttt{PINATA\_JWT} is required for production pins; when absent, a development mock mode generates deterministic local CIDs so UI flows remain testable without external services.
Benchmarks intended for publication (Section~\ref{sec:eval-b}) will use real Pinata pins only.

\subsection{Configuration and Deployment}
Frontend environment variables include \texttt{NEXT\_PUBLIC\_HALALCHAIN\_ADDRESS}, \texttt{NEXT\_PUBLIC\_CHAIN\_ID}, \texttt{NEXT\_PUBLIC\_SEPOLIA\_RPC\_URL}, and \texttt{NEXT\_PUBLIC\_DEMO\_URL} for QR base URLs.
The consumer dashboard is deployed on Vercel (\url{https://web-six-ivory-36.vercel.app/}); contract deployment on Ethereum Sepolia or Base Sepolia requires \texttt{DEPLOYER\_PRIVATE\_KEY} and faucet-funded ETH (Appendix~\ref{app:repro}).
