\section{Evaluation}
\label{sec:eval}

We structure evaluation into two tiers aligned with deployment maturity: (VI-A) qualitative analysis and functional validation available on local Hardhat networks today; and (VI-B) quantitative Base Sepolia benchmarks pending funded testnet deployment.
This separation avoids conflating developer-machine measurements with public network costs~\cite{casino2021systematic}.

\subsection{Qualitative Analysis and Functional Validation}
\label{sec:eval-a}

\subsubsection{Security Properties}
Table~\ref{tab:security} maps design requirements from Section~\ref{sec:prelim} to observable protocol behaviors validated by unit tests and manual scenario runs.

\begin{table}[t]
  \caption{Security properties and validation method.}
  \label{tab:security}
  \centering
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \begin{tabularx}{\columnwidth}{@{}>{\raggedright\arraybackslash}p{0.34\columnwidth}X@{}}
    \toprule
    \textbf{Property} & \textbf{Validation} \\
    \midrule
    Immutability of rejection & \texttt{verifyBatch} on Rejected reverts \\
    Revision traceability & \texttt{parentBatchId} preserved; parent reason readable \\
    Role separation & Non-producer cannot register; non-auditor cannot verify \\
    Batch identity & Invalid \texttt{batchId} reverts on all mutators \\
    Revocation integrity & Only Verified $\rightarrow$ Revoked permitted \\
    \bottomrule
  \end{tabularx}
\end{table}

\paragraph{Immutability.}
Once \texttt{rejectBatch} executes, \texttt{verifyBatch} on the same identifier reverts with \texttt{InvalidStatus}.
Scenario~C (below) confirms that repeated verification attempts leave on-chain state unchanged---supporting \emph{Amanah}.

\paragraph{Revision traceability.}
\texttt{registerRevision} creates a new \texttt{batchId} while storing \texttt{parentBatchId}.
Consumers inspecting a verified revision can navigate to the rejected parent and read its \texttt{rejectReason} and original \texttt{ipfsCid}, supporting audit transparency.

\paragraph{Access control.}
OpenZeppelin \texttt{onlyRole} modifiers gate mutating functions.
Tests assert that unauthorized addresses cannot register, verify, reject, or revoke, addressing basic privilege escalation concerns in the v1 prototype.

\subsubsection{Functional Scenarios}
Three end-to-end scenarios from \texttt{docs/EVALUATION.md} were executed on a local Hardhat network with deployed contracts and the Next.js frontend:

\paragraph{Scenario A (happy path).}
Producer registers batch with IPFS documents $\rightarrow$ auditor verifies $\rightarrow$ consumer page displays Verified status and document links.
Validates the primary UMKM-to-consumer trust path.

\paragraph{Scenario B (rejection and revision).}
Producer registers batch~1 $\rightarrow$ auditor rejects with reason $\rightarrow$ consumer page shows Rejected $\rightarrow$ producer submits \texttt{registerRevision} creating batch~2 $\rightarrow$ auditor verifies batch~2 $\rightarrow$ consumer sees Verified with parent link to batch~1.
Validates revision FSM and UI linkage.

\paragraph{Scenario C (tamper resistance).}
After rejection of batch~1, a direct \texttt{verifyBatch(1)} call reverts.
Subsequent \texttt{getBatch(1)} still returns \texttt{Rejected} with the original reason.
Demonstrates that audit outcomes cannot be silently upgraded without a new batch identifier.

Local validation confirms protocol correctness and UI integration; it does not substitute for public testnet latency, fee volatility, or Pinata production SLAs.

\subsubsection{Development-Network Gas Observations}
Hardhat localhost receipts provide \emph{preliminary} gas usage for contract sizing (Table~\ref{tab:gas-local}).
These figures are \textbf{not} reported as Base Sepolia costs---opcode pricing and L1 data fees differ on rollups~\cite{optimism2024spec}.

\begin{table}[t]
  \caption{Preliminary gas on Hardhat localhost (development only; not Base Sepolia).}
  \label{tab:gas-local}
  \centering
  \footnotesize
  \begin{tabularx}{\columnwidth}{@{}Xr@{}}
    \toprule
    \textbf{Operation} & \textbf{Gas Used} \\
    \midrule
    \texttt{registerBatch} & 155{,}506 \\
    \texttt{verifyBatch} & 79{,}214 \\
    \texttt{rejectBatch} & 103{,}533 \\
    \texttt{registerRevision} & 181{,}835 \\
    \texttt{getBatch} (view) & 0 \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Quantitative Testnet Benchmarks}
\label{sec:eval-b}

Table~\ref{tab:gas} will report gas usage, effective L2 fees, and end-to-end latency measured on Base Sepolia (chain ID 84532) after deployment with faucet-funded accounts.
\textbf{No fabricated testnet values are reported;} cells marked ``TBD'' await on-chain transaction receipts from \texttt{npm run evaluate:gas}.

\begin{table}[t]
  \caption{Gas and cost metrics on Base Sepolia (to be completed after deployment).}
  \label{tab:gas}
  \centering
  \footnotesize
  \begin{tabularx}{\columnwidth}{@{}Xrr@{}}
    \toprule
    \textbf{Operation} & \textbf{Gas Used} & \textbf{Cost (ETH)} \\
    \midrule
    \texttt{registerBatch} & TBD & TBD \\
    \texttt{verifyBatch} & TBD & TBD \\
    \texttt{rejectBatch} & TBD & TBD \\
    \texttt{registerRevision} & TBD & TBD \\
    \texttt{getBatch} (RPC read) & 0 & 0 \\
    \bottomrule
  \end{tabularx}
\end{table}

Planned metrics additionally include: block confirmation time (registration $\rightarrow$ verification), IPFS upload duration via Pinata, and consumer page time-to-first-byte on Vercel hosting.
Run \texttt{npm run evaluate:gas} in \texttt{contracts/} after Base Sepolia deployment (Appendix~\ref{app:repro}).
